Explanatory Data Analysis of Trade Statistics of Turkey

### 2018-12-23 Final/Final20181223.Rmd
.

---
title: "Explanatory Data Analysis of Trade Statistics of Turkey"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    df_print: paged
    number_sections: no
    theme: journal
    toc: yes
date: "December 2018"
subtitle: MEF University - BDA 503 Essentials of Data Analytics - Fall 2018 - Group
  Fou\*R* Term Project
---

@@

## K-Means Clustering

```{r}
product_year<-export_Product_Group %>%
filter(Export_Value > 0) %>%
group_by(Product_Group_Name,year)%>%
summarise(Adet=n(), Toplam = sum(Export_Value))

@@

product_Cluster2 <- scale(product_Cluster2)
head(product_Cluster2)

@@

#Determine optimal cluster with Elbow Method
set.seed(11)
fviz_nbclust(product_Cluster2, kmeans, method = "wss")

#Determine Optimal Cluster with Silhoutte Method
fviz_nbclust(product_Cluster2, kmeans, method = "silhouette")

#Determine Optimal Cluster with Gap Statistics Method
set.seed(12)
gap_stat <- clusGap(product_Cluster2, FUN = kmeans, nstart = 25, K.max = 10, B = 50)
fviz_gap_stat(gap_stat)

# Compute k-means with 4 clusters 
k2 <- kmeans(product_Cluster2, centers = 4, nstart = 25)
print (k2)
fviz_cluster(k2, data = product_Cluster2)

"cluster: A vector of integers (from 1:k) indicating the cluster to which each point is allocated.
centers: A matrix of cluster centers.
totss: The total sum of squares.
withinss: Vector of within-cluster sum of squares, one component per cluster.
tot.withinss: Total within-cluster sum of squares, i.e. sum(withinss).
betweenss: The between-cluster sum of squares, i.e. $totss-tot.withinss$.
size: The number of points in each cluster."

@@

@@

cvv

@@

>>>>

---
title: "Explanatory Data Analysis of Trade Statistics of Turkey"
output:
  html_document:
    code_folding: hide
    df_print: paged
    number_sections: no
    theme: journal
    toc: yes
  pdf_document:
    toc: yes
date: "December 2018"
subtitle: MEF University - BDA 503 Essentials of Data Analytics - Fall 2018 - Group
  Fou\*R* Term Project
---

@@

## K-Means Clustering for Value of Turkey Exports on Products

In order to see how products, which is exported to Turkey, cluster, we implemented k-means clustering method after scaling the values

To determine optimal cluster number, we checked Elbow, Silhoutte and Gap Statistics methods and also we draw distance matrix for products according to export values for 5 years.

```{r, fig.align="center"}
product_year<-export_Product_Group %>%
filter(Export_Value > 0) %>%
group_by(Product_Group_Name,year)%>%
summarise(Adet=n(), Toplam = sum(Export_Value))

@@

product_Cluster2 <- scale(product_Cluster2)

#head(product_Cluster2)

@@

#Determine optimal cluster with Elbow Method
set.seed(11)
Elbow <- fviz_nbclust(product_Cluster2, kmeans, method = "wss")

#Determine Optimal Cluster with Silhoutte Method
Silhoutte <- fviz_nbclust(product_Cluster2, kmeans, method = "silhouette")

#Determine Optimal Cluster with Gap Statistics Method
set.seed(12)
gap_stat <- clusGap(product_Cluster2, FUN = kmeans, nstart = 25, K.max = 10, B = 50)
GapStat <- fviz_gap_stat(gap_stat)

grid.arrange(Elbow, Silhoutte, GapStat, nrow = 1)

@@

It looks like it is better to cluster the data into 2, 3 or 4. So we computed the k-means with 2, 3 and 4 cluster as seen below.

# Compute k-means with 2,3 and 4 clusters 
k1 <- kmeans(product_Cluster2, centers = 2, nstart = 25)
k2 <- kmeans(product_Cluster2, centers = 3, nstart = 25)
k3 <- kmeans(product_Cluster2, centers = 4, nstart = 25)

#print (k3)

k1p <- fviz_cluster(k1, data = product_Cluster2, geom = "point") + ggtitle("k = 2")
k2p <- fviz_cluster(k2, data = product_Cluster2, geom = "point") + ggtitle("k = 3")
k3p <- fviz_cluster(k3, data = product_Cluster2, geom = "point") + ggtitle("k = 4")

grid.arrange(k1p, k2p, k3p, nrow = 3)

fviz_cluster(k3, data = product_Cluster2) + ggtitle("k = 4")

#cluster: A vector of integers (from 1:k) indicating the cluster to which each point is allocated.
#centers: A matrix of cluster centers.
#totss: The total sum of squares.
#withinss: Vector of within-cluster sum of squares, one component per cluster.
#tot.withinss: Total within-cluster sum of squares, i.e. sum(withinss).
#betweenss: The between-cluster sum of squares, i.e. $totss-tot.withinss$.
#size: The number of points in each cluster."

With 4 clusters, it is seen that;

* Vehicles other than railway can be classified as a group - the export products we spent too much money

* Machinery, mechanical appliances, electronical machinery and equipments, clothing is another classification - top2 products in terms of money spent

* Third class consist of iron, steel, pearls, ores etc. and this class slightly differs from 4th group in terms of money spent

@@

As a result of descriptive analysis of Turkey's trade with G7 countries besides India, China and Russia, we noted that;

* Turkeys’ biggest trade partners are China, Germany, Russia and US at import side while they are Germany, UK, Italy and USA at export side. 
* Germany and USA are the 2 top countries seen in both directions of trade.
* Generally we have fluctuated trade trend with Russia.
* Turkey's trade deficit analyisis gives the results : 
  + UK gives trade surplus for all years analysed 
  + China, Russia, Germany and USA have the most significant effect on trade deficit
  + Trade deficit has negatively correlated with USD/TL rate between 2013 and 2016 but there is no relation BIST100 index.
  + Turkey's trade deficit comes from China, Russia, Germany countires and most of deficit comes from products below :
      + Electrical machinery, equipments and machinery, mechanical appliances
      + Mineral fuels, mineral oils and products of their distillation constitute 
      + Vehicles other than railway and mechanical and electronical machinery and equipments
* As a result of clustering exported products;
  + Vehicles other than railway is the group we spent the most money
  + Second group consist of Machinery, mechanical appliances, electronical machinery- equipments and clothing while third one includes iron, steel, pearls, ores etc.

@@

[UC Business Analytics R Programming Guide](https://uc-r.github.io/kmeans_clustering#silo)
